Implications of afforestation policies for global markets

### 2023-08-07 paper.Rmd
Update paper.Rmd

The global gains that can be realized from afforestation, however, are uncertain, with broad estimates of cumulative sequestration ranging from 42 gigatonnes to 700 gigatonnes by the end of the century [@veldman2019]  [@bastin201]  [@Humpenöder2014]  [@sathaye2006]  [@doelman2020]. In addition, the amount of land conversion that would be required also varies greatly, with estimates ranging from 0.3 billion hectares to 2.8 billion hectares [@veldman2019]  [@bastin201]  [@Humpenöder2014]  [@doelman2020]. In many cases, these results suggest massive losses in the amount of pasture and cropland that would be available. This puts into question the feasibility of attaining these results, as almost 90 percent of deforestation worldwide is undertaken to expand the amount of agricultural land [@fao2021]. There are many reasons for this. For example, as nations develop and disposable incomes rise, consumers often change to more meat-based diets. Population growth is also a driver and when it is especially high, which is the case at present throughout most of Africa, this also puts pressure on policymakers to allow greater conversion of forestland into farmland. Still, previous research has suggested that substantive gains in the carbon sink (+23.8PgCo2e/yr) can be made without sacrificing agricultural land [@griscom2017].

Still, the manner in which these various and oftentimes competing concerns are addressed will have significant impacts on the global community.

Moreover, afforestation appears not to provide the same benefits in all regions of the world. In particular, boreal forests appear to provide the smallest net benefits, as the carbon sequestrations benefits of conifer forests are mitigated by the reduction in albedo in winter [@mykleby2017] [@li2015]. Coniferous forests have also been shown to increase soil organic carbon less than broadleaf forests [@guo2021]. With tropical forests, a strong positive result is derived from afforestation, as the carbon sequestration benefit is complemented by a cooling impact due to changes in both albedo and evapotranspiration, and with temperate forests, the impact of afforestation is also shown to be positive, albeit less strongly so due to a winter warming effect [@li2015].

@@

This project will assess the impact of various likely scenarios by melding together two widely accepted global models, the Global Trade Analysis Project (GTAP) computable general equilibrium model [@aguilar2022] and the Global Timber Model (GTM) [@sohngen2001]. The GTAP model has traditionally been used to analyze the impact of various shocks on things like agricultural prices and trade volumes. However, the standard model does not well account for land-use changes, and the GTAP Land Use and Land Cover (LULC) database has limitations due to its infrequent updates. The GTM is quite helpful in getting around these issues, as it is a dynamic optimization model that can examine the impact of afforestation or avoided deforestation on the carbon sink. By joining these two models and passing outputs back and forth between the models, any limitations of either model can be minimized and more robust results, therefore, can be attained.

A transition from deforestation to reforestation, in general, induces the displacement of land use abroad despite the uncertain roles of international trade due to the heterogeneity of regions/countries [@meyfroidt2010] [@lambin2011] DeFries et al. [@defries2010] found a strong correlation between forest loss and exports of agricultural products. On the other hand, Turner et al. [@turner2005] investigated the role of trade on forests in the context of the Free Trade Area of the Americas (FTAA) using the spatial partial-equilibrium Global Forest Products Model and found that FTAA has a positive impact on forest resources.

The Land Use and Land Cover Database within the GTAP framework builds upon the global land cover and land use data, incorporating forestry remote-sensing products [@baldos2012]. Using globally spatially explicit land use data, Global Trade Analysis Project Agro-Ecological Zone (GTAP-AEZ) was introduced in order to enhance the GTAP model by incorporating intra- and inter-regional land and land-based greenhouse emissions heterogeneity [@hertel2008]. It includes competition among different crops, grazing, and forest-based uses within agro-ecological zones (AEZs).

The dynamic GTAP model, supported by the GTAP-AEZ database, considers general equilibrium impacts. Land market effects, identified as significant in driving results by Stevenson et al. [@stevenson2013], are taken into account. The GTAP-AEZ database has been utilized to evaluate the impacts of crop intensification on land use, highlighting the importance of global research and development investment as a strategy to alleviate pressure on natural ecosystems. However, it is important to recognize that global aggregates may mask localized shifts that can have implications for ecologically significant areas [@byerlee2014].

Steinbuks and Hertel [@steinbuks2012] developed the FABLE model (Forestry, Agriculture, Biofuels, Land Use, and Environment) as a global partial equilibrium model to analyze optimal land use within a context of increasing demand for food, bioenergy, forest products, and non-provisioning ecosystem services, while also meeting greenhouse gas targets. The model encompasses nine sectors, including agriculture, livestock, food processing, biofuels, energy, forest, timber processing, and ecosystem services. While FABLE is a powerful tool for examining the optimal trajectory of various land uses under specific demand assumptions, it primarily focuses on specific sectors of the economy, neglecting general equilibrium effects and socio-economic or household welfare impacts [@steinbuks2012] [@hertel2017].

The KLUM@GTAP framework integrates the Kleines Land Use Model (KLUM) with an extended version of GTAP called GTAP-EFL to evaluate the impacts of climate change on cropland allocation. GTAP-EFL separates energy factors from intermediate inputs and incorporates them into capital, while also considering CO2 emissions. KLUM, a global agricultural land-use model, connects the economy to global crop allocation to maximize producer returns under specific risk assumptions. KLUM@GTAP substitutes the land allocation mechanism within GTAP-EFL by utilizing regionally aggregated area changes in cropland determined by KLUM to update cropland shares in GTAP-EFL [@ronneberger2009].

@@

# References

>>>>

The global gains that can be realized from afforestation, however, are uncertain, with broad estimates of cumulative sequestration ranging from 42 gigatonnes to 700 gigatonnes by the end of the century [@veldman2019]  [@bastin201]  [@Humpenöder2014]  [@sathaye2006]  [@doelman2020]. In addition, the amount of land conversion that would be required also varies greatly, with estimates ranging from 0.3 billion hectares to 2.8 billion hectares [@veldman2019]  [@bastin201]  [@Humpenöder2014]  [@doelman2020]. In many cases, these results suggest massive losses in the amount of pasture and cropland that would be available. This puts into question the feasibility of attaining these results, as almost 90 percent of deforestation worldwide is undertaken to expand the amount of agricultural land [@fao2021]. There are many reasons for this. For example, as nations develop and disposable incomes rise, consumers often change to more meat-based diets. Population growth is also a driver and when it is especially high, which is the case at present throughout most of Africa, this also puts pressure on policymakers to allow greater conversion of forestland into farmland.

Land-use decisions in one countr also tend to have impacts that spill over into many others [@meyfroidt2010] [@lambin2011], and therefore, cumulative change needs to be assessed with global-scale models.  Steinbuks and Hertel [@steinbuks2012] developed the FABLE model (Forestry, Agriculture, Biofuels, Land Use, and Environment) as a global partial equilibrium model to analyze optimal land use within a context of increasing demand for food, bioenergy, forest products, and non-provisioning ecosystem services, while also meeting greenhouse gas targets. The model encompasses nine sectors, including agriculture, livestock, food processing, biofuels, energy, forest, timber processing, and ecosystem services. While FABLE is a powerful tool for examining the optimal trajectory of various land uses under specific demand assumptions, it primarily focuses on specific sectors of the economy, neglecting general equilibrium effects and socio-economic or household welfare impacts [@steinbuks2012] [@hertel2017]. There have been attempts to use the Global Trade Analysis Project (GTAP) computable general equilibrium model to address such questions, but the standard model does not well account for land-use changes. The Land Use and Land Cover Database within the GTAP framework, which incorporated forestry remote-sensing products, was thus created [@baldos2012]. It was strengthened as part of the GTAP Agro-Ecological Zone (GTAP-AEZ) model, with the use of spatially explicit global land use data and through the incorporation of intra- and inter-regional land and land-based greenhouse emissions heterogeneity [@hertel2008]. It also includes competition among different crops, grazing, and forest-based uses, but it has limitations due to its infrequent updates. The KLUM@GTAP framework integrates the Kleines Land Use Model (KLUM) with an extended yet static version of GTAP called GTAP-EFL to evaluate the impacts of climate change on cropland allocation. GTAP-EFL separates energy factors from intermediate inputs and incorporates them into capital, while also considering CO2 emissions. KLUM, a global agricultural land-use model, connects the economy to global crop allocation to maximize producer returns under specific risk assumptions. KLUM@GTAP substitutes the land allocation mechanism within GTAP-EFL by utilizing regionally aggregated area changes in cropland determined by KLUM to update cropland shares in GTAP-EFL [@ronneberger2009].

More recent modeling attempts of land-use change have begun to move away from comparative static analysis. For example, the dynamic GTAP model, supported by the GTAP-AEZ database, considers land market effects, which were identified as significant in driving results by Stevenson et al. [@stevenson2013], and has been utilized to evaluate the impacts of crop intensification on land use. It, however, has been critiqued as it global aggregates may mask localized shifts that can have implications for ecologically significant areas [@byerlee2014]. This project will attempt to get around these concerns and assess the impact of various likely afforestation scenarios by melding GTAP with the Global Timber Model (GTM) [@sohngen2001]. The GTM is seen as being quite helpful in getting around these issues, as it is a dynamic optimization model that can examine the impact of afforestation or avoided deforestation on the carbon sink. By joining these two models and passing outputs back and forth between the models, any limitations of either model can be minimized and more robust results, therefore, can be attained.

Still, previous research has suggested that substantive gains in the carbon sink (+23.8PgCo2e/yr) can be made without sacrificing agricultural land [@griscom2017].

Faced with competing land-use demands, afforestation efforts need to be targeted to the regions of the world where they will provide the most benefits. Afforestation in boreal zones has been shown to provide the smallest net benefits, as the carbon sequestrations benefits of conifer forests are mitigated by the reduction in albedo in winter [@mykleby2017] [@li2015]. Coniferous forests have also been shown to increase soil organic carbon less than broadleaf forests [@guo2021]. Conversely, tropical forests have been linked with a strong positive result from afforestation, as the carbon sequestration benefit is complemented by a cooling impact due to changes in both albedo and evapotranspiration, and with temperate forests, the impact of afforestation is also shown to be positive, albeit less strongly so due to a winter warming effect [@li2015].

@@

@@

DeFries et al. [@defries2010] found a strong correlation between forest loss and exports of agricultural products. On the other hand, Turner et al. [@turner2005] investigated the role of trade on forests in the context of the Free Trade Area of the Americas (FTAA) using the spatial partial-equilibrium Global Forest Products Model and found that FTAA has a positive impact on forest resources.
# References
